### 2011-07-16 cover_letter.tex
Addressed concerns regarding missing part in the algorithm

\documentclass[11pt]{article}
\usepackage[inner=2cm,outer=2cm, top=2cm, bottom=2cm]{geometry}               % See geometry.pdf to learn the layout options. There are lots.

\usepackage{pseudocode}
\usepackage{url}

\title{Responses to reviewers' comments and list of modifications done\\ on paper SORO226\\
\LARGE{\textbf{Grounding the interaction: anchoring situated discourse in everyday human-robot interaction}}}
\author{S\'everin Lemaignan \and Raquel Ros \and E. Akin Sisbot \and Rachid Alami \and Michael Beetz}

\begin{document}

\maketitle

The authors would like to thank the reviewers for their constructive feedback that helped us to strengthen the paper.

Below follow our answers to the reviewersÕ comments and suggestions. Each answer also indicates the changes made in the paper to address the reviewerÕs concern. An important number of language corrections are made, and the paper is proof-read by a native English speaker.

%We are very grateful to the reviewers for having accepted to review our 
%work, for the care they took in examining it and their resulting 
%comments which have been appreciated in order to improve the quality of
%the paper.

%Several reviewers noted linguistic issues and terminology/typography 
%inconsistencies. We have done our best to carefully proof-read the article,
%with the help of a native English speaker.

%In the following three sections, we address separately each of
%the concerns of our three reviewers.

\section{Reviewer 1}
%\section{Comments of the first reviewer}

\subsection{How robust is the approach with regard to perceptual recognition
errors?}

We are aware that perception is not robust and therefore one of our next steps
is to include a probabilistic layer to the approach (as already mentioned in
the discussion section). We believe though that a robot (specially interacting
with people) should be allowed to make mistakes due to perception noise, or not
to be sure about something (for instance, "I think that..."), which represents
the uncertainty the robot has about a fact. In the meantime, to overcome
perception errors, the robot provides visual feedback (looking at objects) to
the user when possible and/or verifies the information with the user whenever
is needed (for instance, "do you mean the box\_01"?).

\subsection{Does the cognitive model enable a meaningful interaction with a
real user in a real-world setting?}

Due to the perception problems and computational delays, fully realistic
settings are not yet possible.  The interaction based only on a static setting,
i.e. objects don't move, the interaction with a real user is accurate enough.
We have developed some experiments, where the user interacts with the system
only through a keyboard (no gestures are included) asking questions, providing
information, etc. and the outcome is quite promising. However, for dynamic
settings where the user can point at things, move the objects around, etc., the
interaction is not fast enough and the user should perform slow movements for
the robot to create a consistent state of world.

A recent video showing the current state of the interaction can be viewed
here: \url{http://www.youtube.com/watch?v=IODx50uV_k4}

\subsection{Rework of the introduction}

Reviewer comment: ``The paper starts off with a discussion
of Austins speech act theory, including locutionary act, illocutionary force
and perlocutionary act. As the concept of speech acts is not really picked up
in the remainder of the paper, I do not see the point in starting with such a
discussion. It would be more appropriate to motivate how a robot could benefit
from a symbolic knowledge base etc.''

Austin's \emph{speech acts} were actually one of the starting point of our
analysis of dialogue understanding, and we think it is worthwhile to briefly mention
them in the introduction, as one of the theoritical ground of our work.

The reviewer however rightfully notes that \emph{perlocutionnary acts}, while mentioned, are not
actually used further. We moved this paragraph to a simple footnote.

\section{Reviewer 2}
% \section{Comments of the second reviewer}

\begin{itemize}
\item All reviewer comments regarding citations have been addressed, following reviewer's suggestions,
\item The reviewer's remark on human tendency to explain in a consistent way the world has been added
\item Language issues have been addressed as well.
\end{itemize}

\section{Reviewer 3}
%\section{Comments of the third reviewer}

\subsection{``Explain the relation of the presented approach to the Open World
Assumption''}

The Open World assumption mainly allows us to model that we may not know
if a fact is true or not.

One very common example is the {\tt knowsAbout} predicate. This predicate
defines if some agent knows about some object.

It is important (especially for dialog) to be able to equally model that an agent knows or 
does not know about the object, or that the robot simply does not know if 
the human knows about it.

The Open World Assumption has important consequences on the reasoning
abilities of the system, since it prevents the classical "negation by 
failure" and leads to tighter modelling requirement to locally "close the
world" (for instance, the {\tt Color} class is defined as the equivalent of
the union of the 10 or so color instances we consider).

\subsection{``Present a systematic overview of the formalisms and
representations used (OWL, RDF, DL, ...), and choose one consistent syntax for
listing the examples''}

The paragraph 2.2 has been reorganized to clarify the vocabulary and better
define the various acronyms.

\subsection{``The authors are encouraged to explain the notion of curiosity
with respect to the Dora robot by Hawes et al. (2011, ICRA) and Hanheide et al.
(2011, IJCAI)''}

A mention to the Dora robot has been added in section 2.1.

\subsection{``Give examples of the alignment of ORO with OpenCyc''}

Numerous predicates in {\tt commonsense.oro.owl} are defined in the 
OpenCyc namespace: {\tt farFrom, near, hold-Underspecified, objectFoundInLocation}.

Similarly, most of the upper classes in the ontology belong to OpenCyc 
(like {\tt SpatialThing, PartiallyTangible, TemporalThing}...).

Overall, about 50 entites (on the >180 defined in {\tt commonsense.oro.owl}) 
come from OpenCyc.

Since most of the upper classes are OpenCyc concepts, this should 
guarantee to a certain extend the alignment of our ontology with other ontologies.

\subsection{Typography and notation consistency}

Typography has been throughly revised for better consistency. Notation variants have been
made explicit (for instance, section 2.4).

\subsection{``p.6, II.E. Explain "multi-lingual support"''}

The "multi-lingual support" has been reworded:

``fast concept lookup, with possible multi-lingual support (through the
{\tt @lang} XML annotation, labels of concept can be translated, and
specific translations can be queried for)''

\subsection{``Points At: please describe how you manage the critical aspect of
timing''}

PointsAt relation is computed continuously during the interaction by 
selecting the objects that "might" be pointed at by the human. The outcome
of the computation (facts like "human1 pointsAt bottle1") are then
continously updated in the knowledge base.

In the current system, the decision on whether the human is pointing at an 
object in a given time is left to the reasoning on the dialog or on the human gestures:
when the dialogue module analyses a sentence like "give me this", it tries to
resolve the atom "this" by querying the ontology.

Both sitatuation assessement and dialogue processing chains induce 
small delays, but these are roughly the same (below 200ms), and 
compensate for each other.

\subsection{``p.8, III.B.Location according to an object: the authors are
kindly referred to the work done by, e.g., K. Sj\"o\"o or J. Kelleher''}

Kelleher's work, as well as Blisard's, are now correctly cited in the section.

\subsection{``p.8, III.B. "creation of imaginary objects...hinted by verbal
assertions..." -- this phenomenon is called presupposition accommodation''}

We thank the reviewer for this input that has been added to the article.

\subsection{``p.9, Alg. IV.1: the crucial bits are left out; please explain
"GenerateDescription", "Ontology.Find", "Ontology.CheckEquivalent", and
"Discrimination"''}

{\tt GenerateDescription} translates the parsing tree of each nominal group
into statements. A set of rules (like the one for adjectives mentioned 
in footnote, section 5.1) help with this task.

The detailled algorithm is provided in appendix to this letter.
We would be happy to add it to the paper as well if the editor want to consider it.

{\tt Ontlogy.Find} and {\tt Ontlogy.CheckEquivalent} are simple, atomic calls to
the ontology server. The first one send a SPARQL query build from the set of 
partial statement build at the {\tt GenerateDescription} step, the second
one check for each candidate pair if {\tt candidate1 owl:sameAs candidate2}.
This check uses the classified knowledge model.

The {\tt Discrimation} algorithm is explained in details in Ros2010. The
reference has been made more explicit.

\subsection{``how does the system deal with restrictive information vs.
attributive information expressed by adjectives (e.g., in "the yellow banana is
big")?''}

\subsection{``Mention work on Generation and Resolution of Referring Expressions''}

Thanks for pointing out the work that has been done in this research community.
Even if our contribution focuses more on the grounding of concepts from 
multi-modal knowledge source, these contribution are of definitive 
interest for us.

A reference to Zender's work has been added to the \emph{Related work} 
section, along with a reference to Dale and Reiter paper from 1995.

\subsection{``Present an initial overview of the
robot hardware and the common software architecture used in the systems''}

Paragraph added in the section 6 introduction.

\subsection{``VI. which university in Munich? LMU or TUM?''}

The study was conducted with TUM in Munich. This has been corrected in 
the paper as well.

\subsection{``VI. please present the anecdote of unexpected behavior elsewhere
or give some more context''}

\section{Appendix}

\subsection{GenerateDescription algorithms}
\label{GenerateDescription}
\begin{pseudocode}[ruled]{GenerateDescription}{group}
\label{algo|GenerateDescription}

\PROCEDURE{GenerateDescription}{group} 
   noun \GETS \CALL{GetNoun}{group} \\ 
   \IF \CALL{Ontology.Lookup}{noun} \in (Instances) \THEN
   		\BEGIN
		id \GETS \CALL{Ontology.lookup}{noun}\\	
		\RETURN {\mathcal{D} + \{ *\ {\tt sameAs}\ <id> \}}\\
		\END
   \ELSE
    	\mathcal{D} = \mathcal{D} + \{ *\ {\tt type}\ <noun>\} \\

   \\
   det \GETS \CALL{GetDeterminant}{group} \\
   \IF det \in {\mbox(possessives)} \THEN
       \mathcal{D} = \mathcal{D} + \{ *\ {\tt isRelatedTo}\ <possessor>\} \\

    \IF det \in {\mbox(demonstratives)} \THEN
        \BEGIN
        \IF \CALL{Ontology.Check}{\{<currentSpeaker>\ {\tt focusesOn}\ *\}} \THEN 
            \mathcal{D} = \mathcal{D} + \{<currentSpeaker>\ {\tt focusesOn}\ *\}
        \ELSE
            \mathcal{D} = \mathcal{D} + \CALL{AnaphoricMatching}{} \\
        \END \\
   \\
   adjs \GETS \CALL{GetAdjectives}{group} \\
   \FOREACH adj \in adjs \DO
   	\BEGIN
   		\IF adj == <other> \THEN 
   			\BEGIN
   			id \GETS \CALL{History.GetMatchingGroup}{group} \STMTNUM{8em}{st.history}\\
   			\mathcal{D} = \mathcal{D} + \{ *\ {\tt differentFrom}\ <id> \}\\
   			\RETURN{D}\\
			\END   		
   		\ELSE
	     	\mathcal{D} = \mathcal{D} + \{ *\ {\tt hasFeature}\ <adj>\} \\
    \END\\

   \\  
   nounComplements \GETS \CALL{GetNounComplements}{group} \\
   \FOREACH nouncmpl \in nounComplements \DO
     \mathcal{D} = \mathcal{D} + {\CALL{GenerateDescription}{nouncmpl}}\\

   \\  
   relativeClauses \GETS \CALL{GetSubordinateRelativeClauses}{group} \\
   \FOREACH relative \in relativeClauses \DO
   	\BEGIN
   	 \mathcal{G} \GETS \CALL{GetNominalGroups}{relative} \\
   	 \FOREACH g \in \mathcal{G} \DO
     	\mathcal{D} = \mathcal{D} + {\CALL{GenerateDescription}{g}}
    \END\\

   \\
   \RETURN{\mathcal{D}} 
\ENDPROCEDURE
\end{pseudocode}

The method called at (\ref{st.history}) consists in looking through sentences 
that have been stored in the conversation history, then extracting their nominal
group in order to retrieve the identifier of the most recently mentioned 
concept that holds the same characteristics as the nominal group that is being 
processed (algorithm~\ref{algo|History}).

\begin{pseudocode}[ruled]{History.GetMatchingGroup}{group}
\label{algo|History}
\PROCEDURE{History.GetMatchingGroup}{group}
\COMMENT{Extract Nominal group from sentences stored in the history}\\
\mathcal{H} \GETS \CALL{History.GetAllNominalGoup}{}\\
\COMMENT{Generate description of the nominal group that is being processed.} \\
\COMMENT{The adjective  "other" is to be removed before calling this routine} \\
	\mathcal{G} \GETS \CALL{GenerateDescription}{group} \\

	candidates \GETS \mathcal{H} \cap \mathcal{G}\\
	\IF \left|{candidates}\right| = 0 \THEN
    \BEGIN
      \OUTPUT{\mbox{Couldn't find another object with the same characteristics!}} \\
      \EXIT \\
    \END
   \ELSEIF \left|{candidates}\right| = 1 \THEN
      id \GETS candidates[0]
   \ELSE
   	  id \GETS \CALL{Discrimination}{candidates}\\
   \RETURN{id}
\ENDPROCEDURE
\end{pseudocode}

\end{document}